\papertitle

### 2024-09-13 src/tex/ms.tex
Update src/tex/ms.tex

\begin{figure}
    \centering
    \resizebox{\columnwidth}{!}{\includegraphics*{figure1.pdf}}
    \caption{
        Example distributions of estimated self-diffusion coefficients, $\Dest$, calculated using (a) ordinary least squares (OLS), (b) weighted least squares (WLS), and (c) generalised least squares (GLS),
        from MSD data from \num{4096} individual simulations of \num{128} particles undergoing a \SI{128}{step} 3D lattice random walk, with a step size chosen so that the true diffusion coefficient $\D = 1$.
        In each panel, the grey curve shows the best-fit normal distribution for the simulation data, the upper horizontal bar shows the standard deviation of this distribution, and the lower horizontal bar shows the average estimated standard deviation given by the analytical expression for $\sigma[\prob{\Dest}]$ for each regression method.}
    \label{fig:glswlsols}
    \script{glswlsols.py}
\end{figure}

@@

The estimated variance $\varest{\oMSD}$ can be calculated from a single simulation trajectory, and provides an accurate estimate of the true variance $\var{\oMSD}$~\footnote{A common approach to estimating the variance of the mean of time-correlated data with unknown correlation time is the renormalization group blocking method of Flyvberg and Peterson~\cite{FlyvbjergAndPetersen_JChemPhys1989,Frenkel2023-ah}. In the SI, we compare this blocking method to our direct rescaling method (\cref{equ:variance}) for estimating $\var{\oMSDi}$ from a set of random-walk trajectories. For our example data, direct rescaling performs better (gives more accurate estimates of $\var{\oMSDi}$).}.
To demonstrate this, we performed \num{4096} independent simulations of \num{128} particles undergoing a three-dimensional cubic-lattice random walk of \num{128} steps per particle.
Using data from all \num{4096} simulations, we first compute the true simulation MSD and its variance (\cref{fig:msd}a).
We also compute the MSD and estimated variance using data from a single simulation trajectory (\cref{fig:msd}b), using the scheme described above.
A quantitative comparison between the true MSD variance and the single-trajectory estimated MSD variance is made in \cref{fig:msd}c: the close numerical agreement confirms that \cref{equ:varestMSD} can be used to estimate $\var{\oMSD}$, which can then be used to parameterise the model covariance matrix $\mathbf{\Sigma^\prime}$ via \cref{equ:cvv}.
\begin{figure}
    \centering
    \resizebox{\columnwidth}{!}{\includegraphics*{figure2.pdf}}
    \caption{
        Comparison of the numerical variance in observed MSD from multiple replica simulations and the estimated variance in observed MSD given by rescaling the variance in observed squared displacements (\cref{equ:varestMSD}).
        Panel (a) shows the mean observed MSD from \num{4096} simulations of \num{128} particles undergoing a 3D lattice random walk of \num{128} steps per particle, with error bars of $\pm2\sigma[\oMSDi]$.
        Panel (b) shows the MSD from just one simulation, with error bars of $\pm2\widehat{\sigma}[\oMSDi]$, obtained via \cref{equ:varestMSD}.
        Panel (c) plots the numerical variance against the estimated variance from a single simulation as a function of timestep $i$.
    }
    \label{fig:msd}
    \script{msd.py}
\end{figure}

@@

To illustrate the accuracy of the numerical procedure for deriving the model covariance matrix, $\bm{\Sigma^\prime}$, we present in \cref{fig:covariances} the MSD covariance matrix for \num{4096} random-walk simulations, as described above, at three differing levels of approximation:
\cref{fig:covariances}a shows the numerically converged covariance matrix, $\bm{\Sigma}$, computed using the data from all \num{4096} simulations (\cref{fig:covariances}a); \cref{fig:covariances}b shows the corresponding analytical covariance matrix, $\mathbf{\Sigma^\prime}$, as defined by \cref{equ:cvv} and parametrised using analytical long-time-limit variances $\var{\oMSDi}$; and \cref{fig:covariances}c shows the average estimated matrix obtained by parametrising \cref{equ:cvv} using variances estimated from a single simulation trajectory, then reconditioning, with the average taken over all \num{4096} matrices obtained from the \num{4096} input simulations. 
\begin{figure}
    \centering
    \resizebox{\columnwidth}{!}{\includegraphics*{figure3.pdf}}
    \caption{
        (a) The numerical MSD covariance matrix $\bm{\Sigma}$ calculated using MSD data from \num{4096} simulations of \num{128} particles undergoing a 3D lattice random walk of \num{128} steps per particle.
        (b) The analytical MSD covariance matrix $\bm{\Sigma^\prime}$ (\cref{equ:cvv}), parametrised using analytical long-time limit random-walk variances $\var{\oMSDi}$.
        (c) The MSD covariance matrix obtained applying the numerical scheme described in the main text to each individual random walk simulation, averaged over all \num{4096} such simulations.
        Colour bars in (a--c) show the covariance, $\Sigma\left[\oMSDi, \oMSDj\right]$.
        The off-diagonal panels show difference plots, computed as per-element ratios between pairs of covariance matrices (a--c).
    }
    \label{fig:covariances}
    \script{covariances.py}
\end{figure}

\begin{figure*}[ht!]
    \centering
    \resizebox{\textwidth}{!}{\includegraphics*{figure4.pdf}}
    \caption{
        (a) Observed MSD from a single simulation of 128 particles undergoing a 3D-lattice random walk of 128 steps per particle (dark line).
        The green shading shows the corresponding posterior distribution $\prob{\model|\oMSD}$ of linear models compatible with the observed MSD data $\oMSD$, calculated using the scheme described in the main text.
        The variegated shading indicates compatibility intervals of \SIlist[list-final-separator = {, and }]{1;2;3}{\sigma}$[\prob{\model|\oMSD}]$.
        (b) The marginal posterior distribution $\prob{\Dest|\oMSD}$ obtained from the posterior distribution of linear models in (a).
        The mean of this distribution gives the point estimate $\Dest$ for this simulation input data.
        The blue horizontal bar shows an interval of one standard deviation in $\prob{\Dest|\oMSD}$.
        (c) Probability distribution of point-estimates $\prob{\Dest}$ obtained from \num{4096} individual random-walk simulations.
        Each simulation has been analysed as in (a) and (b) to yield a single corresponding point estimate $\Dest$.
        The grey line shows the distribution of point estimates, $\prob{\Dest_\mathrm{num}}$, obtained using Bayesian regression with a mean vector and numerical covariance matrix derived from the complete dataset of all \num{4096} simulations.
        The pink horizontal bar shows an interval of one standard deviation in $\prob{\Dest}$.
        (d) Probability distribution of estimated variances, $\varest{\Dest}$, for individual random-walk simulations, compared to the true sample variance (pink vertical line) $\var{\Dest}$.
    } 
    \label{fig:random_walk}
    \script{random_walk.py}
\end{figure*}

@@

\begin{figure}[tb]
    \centering
    \resizebox{\columnwidth}{!}{\includegraphics*{figure5.pdf}}
    \caption{
      (a) Probability distribution of point estimates $\prob{\Dest}$ for \num{192} effective simulations of LLZO (orange histogram).
      The grey line shows the distribution $\prob{\Dest_\mathrm{num}}$ obtained using Bayesian regression with the complete LLZO dataset as input.
      The pink bar shows an interval of one standard deviation $\sigma[\prob{\Dest}]$.
      (b) Probability distribution of estimated variances, $\varest{\Dest}$, for individual LLZO effective simulations, compared to the true sample variance (pink vertical line) $\var{\Dest}$.
    }
    \label{fig:diffusion}
    \script{diffusion.py}
\end{figure}

@@

\begin{figure}[htb]
    \centering
    \resizebox{\columnwidth}{!}{\includegraphics*{figure6.pdf}}
    \caption{
      Scaling of $\var{\Dest}$ with simulation size for OLS (pink), WLS (blue), our approximate Bayesian regression method (green), and GLS (orange).
      (a) Scaling versus number of mobile particles, $N_\mathrm{atoms}$.
      (b) Scaling versus total simulation time, $t_\mathrm{max}$.
      Solid lines show fitted power law relationships for each dataset.
      The WLS and GLS data are obtained using numerically determined variances and covariance, respectively, from a set of \num{512} repeat simulations for each combination of $N_\mathrm{atoms}$ and $t_\mathrm{max}$.
    }
    \label{fig:stat_eff}
    \script{stat_eff.py}
\end{figure}

@@

Electronic Supplementary Information (ESI) available: A complete set of analysis/plotting scripts allowing for a fully reproducible and automated analysis workflow, using \textsc{showyourwork}~\cite{luger_showyourwork_2021}, for this work is available at Ref.~\cite{mccluskey_github_2022} under an MIT license.
All raw simulation files are available on Zenodo shared under CC BY-SA 4.0 licences, the random walk simulations and other analysis-linked data~\cite{mccluskey_zenodo_2022} and the LLZO raw simulation trajectories~\cite{coles_llzo_zenodo_2022}.
The method outlined in this work is implemented in the open-source Python package \textsc{kinisi}~\cite{mccluskey_kinisi_2022}, which is available under an MIT license.

>>>>

\begin{figure}
    \centering
    % \resizebox{\columnwidth}{!}{\includegraphics*{figure1.pdf}}
    \includegraphics[width=\columnwidth]{glswlsols.pdf}
    \caption{
        Example distributions of estimated self-diffusion coefficients, $\Dest$, calculated using (a) ordinary least squares (OLS), (b) weighted least squares (WLS), and (c) generalised least squares (GLS),
        from MSD data from \num{4096} individual simulations of \num{128} particles undergoing a \SI{128}{step} 3D lattice random walk, with a step size chosen so that the true diffusion coefficient $\D = 1$.
        In each panel, the grey curve shows the best-fit normal distribution for the simulation data, the upper horizontal bar shows the standard deviation of this distribution, and the lower horizontal bar shows the average estimated standard deviation given by the analytical expression for $\sigma[\prob{\Dest}]$ for each regression method.}
    \label{fig:glswlsols}
    \script{glswlsols.py}
\end{figure}

@@

The estimated variance $\varest{\oMSD}$ can be calculated from a single simulation trajectory, and provides an accurate estimate of the true variance $\var{\oMSD}$~\footnote{A common approach to estimating the variance of the mean of time-correlated data with unknown correlation time is the renormalization group blocking method of Flyvberg and Peterson~\cite{FlyvbjergAndPetersen_JChemPhys1989,Frenkel2023-ah}. In the SI, we compare this blocking method to our direct rescaling method (\cref{equ:variance}) for estimating $\var{\oMSDi}$ from a set of random-walk trajectories. For our example data, direct rescaling performs better (gives more accurate estimates of $\var{\oMSDi}$).}.
To demonstrate this, we performed \num{4096} independent simulations of \num{128} particles undergoing a three-dimensional cubic-lattice random walk of \num{128} steps per particle.
Using data from all \num{4096} simulations, we first compute the true simulation MSD and its variance (\cref{fig:msd}a).
We also compute the MSD and estimated variance using data from a single simulation trajectory (\cref{fig:msd}b), using the scheme described above.
A quantitative comparison between the true MSD variance and the single-trajectory estimated MSD variance is made in \cref{fig:msd}c: the close numerical agreement confirms that \cref{equ:varestMSD} can be used to estimate $\var{\oMSD}$, which can then be used to parameterise the model covariance matrix $\mathbf{\Sigma^\prime}$ via \cref{equ:cvv}.
\begin{figure}
    \centering
    % \resizebox{\columnwidth}{!}{\includegraphics*{figure2.pdf}}
    \includegraphics[width=\columnwidth]{msd.pdf}
    \caption{
        Comparison of the numerical variance in observed MSD from multiple replica simulations and the estimated variance in observed MSD given by rescaling the variance in observed squared displacements (\cref{equ:varestMSD}).
        Panel (a) shows the mean observed MSD from \num{4096} simulations of \num{128} particles undergoing a 3D lattice random walk of \num{128} steps per particle, with error bars of $\pm2\sigma[\oMSDi]$.
        Panel (b) shows the MSD from just one simulation, with error bars of $\pm2\widehat{\sigma}[\oMSDi]$, obtained via \cref{equ:varestMSD}.
        Panel (c) plots the numerical variance against the estimated variance from a single simulation as a function of timestep $i$.
    }
    \label{fig:msd}
    \script{msd.py}
\end{figure}

@@

To illustrate the accuracy of the numerical procedure for deriving the model covariance matrix, $\bm{\Sigma^\prime}$, we present in \cref{fig:covariances} the MSD covariance matrix for \num{4096} random-walk simulations, as described above, at three differing levels of approximation:
\cref{fig:covariances}a shows the numerically converged covariance matrix, $\bm{\Sigma}$, computed using the data from all \num{4096} simulations (\cref{fig:covariances}a); \cref{fig:covariances}b shows the corresponding analytical covariance matrix, $\mathbf{\Sigma^\prime}$, as defined by \cref{equ:cvv} and parametrised using analytical long-time-limit variances $\var{\oMSDi}$; and \cref{fig:covariances}c shows the average estimated matrix obtained by parametrising \cref{equ:cvv} using variances estimated from a single simulation trajectory, then reconditioning, with the average taken over all \num{4096} matrices obtained from the \num{4096} input simulations. 
\begin{figure}
    \centering
    % \resizebox{\columnwidth}{!}{\includegraphics*{figure3.pdf}}
    \includegraphics[width=\columnwidth]{covariances.pdf}
    \caption{
        (a) The numerical MSD covariance matrix $\bm{\Sigma}$ calculated using MSD data from \num{4096} simulations of \num{128} particles undergoing a 3D lattice random walk of \num{128} steps per particle.
        (b) The analytical MSD covariance matrix $\bm{\Sigma^\prime}$ (\cref{equ:cvv}), parametrised using analytical long-time limit random-walk variances $\var{\oMSDi}$.
        (c) The MSD covariance matrix obtained applying the numerical scheme described in the main text to each individual random walk simulation, averaged over all \num{4096} such simulations.
        Colour bars in (a--c) show the covariance, $\Sigma\left[\oMSDi, \oMSDj\right]$.
        The off-diagonal panels show difference plots, computed as per-element ratios between pairs of covariance matrices (a--c).
    }
    \label{fig:covariances}
    \script{covariances.py}
\end{figure}

\begin{figure*}[ht!]
    \centering
    % \resizebox{\textwidth}{!}{\includegraphics*{figure4.pdf}}
    \includegraphics[width=\textwidth]{random_walk.pdf}
    \caption{
        (a) Observed MSD from a single simulation of 128 particles undergoing a 3D-lattice random walk of 128 steps per particle (dark line).
        The green shading shows the corresponding posterior distribution $\prob{\model|\oMSD}$ of linear models compatible with the observed MSD data $\oMSD$, calculated using the scheme described in the main text.
        The variegated shading indicates compatibility intervals of \SIlist[list-final-separator = {, and }]{1;2;3}{\sigma}$[\prob{\model|\oMSD}]$.
        (b) The marginal posterior distribution $\prob{\Dest|\oMSD}$ obtained from the posterior distribution of linear models in (a).
        The mean of this distribution gives the point estimate $\Dest$ for this simulation input data.
        The blue horizontal bar shows an interval of one standard deviation in $\prob{\Dest|\oMSD}$.
        (c) Probability distribution of point-estimates $\prob{\Dest}$ obtained from \num{4096} individual random-walk simulations.
        Each simulation has been analysed as in (a) and (b) to yield a single corresponding point estimate $\Dest$.
        The grey line shows the distribution of point estimates, $\prob{\Dest_\mathrm{num}}$, obtained using Bayesian regression with a mean vector and numerical covariance matrix derived from the complete dataset of all \num{4096} simulations.
        The pink horizontal bar shows an interval of one standard deviation in $\prob{\Dest}$.
        (d) Probability distribution of estimated variances, $\varest{\Dest}$, for individual random-walk simulations, compared to the true sample variance (pink vertical line) $\var{\Dest}$.
    } 
    \label{fig:random_walk}
    \script{random_walk.py}
\end{figure*}

@@

\begin{figure}[tb]
    \centering
    % \resizebox{\columnwidth}{!}{\includegraphics*{figure5.pdf}}
    \includegraphics[width=\columnwidth]{diffusion.pdf}
    \caption{
      (a) Probability distribution of point estimates $\prob{\Dest}$ for \num{192} effective simulations of LLZO (orange histogram).
      The grey line shows the distribution $\prob{\Dest_\mathrm{num}}$ obtained using Bayesian regression with the complete LLZO dataset as input.
      The pink bar shows an interval of one standard deviation $\sigma[\prob{\Dest}]$.
      (b) Probability distribution of estimated variances, $\varest{\Dest}$, for individual LLZO effective simulations, compared to the true sample variance (pink vertical line) $\var{\Dest}$.
    }
    \label{fig:diffusion}
    \script{diffusion.py}
\end{figure}

@@

\begin{figure}[htb]
    \centering
    % \resizebox{\columnwidth}{!}{\includegraphics*{figure6.pdf}}
    \includegraphics[width=\columnwidth]{stat_eff.pdf}
    \caption{
      Scaling of $\var{\Dest}$ with simulation size for OLS (pink), WLS (blue), our approximate Bayesian regression method (green), and GLS (orange).
      (a) Scaling versus number of mobile particles, $N_\mathrm{atoms}$.
      (b) Scaling versus total simulation time, $t_\mathrm{max}$.
      Solid lines show fitted power law relationships for each dataset.
      The WLS and GLS data are obtained using numerically determined variances and covariance, respectively, from a set of \num{512} repeat simulations for each combination of $N_\mathrm{atoms}$ and $t_\mathrm{max}$.
    }
    \label{fig:stat_eff}
    \script{stat_eff.py}
\end{figure}

@@

Electronic Supplementary Information (ESI) available: A complete set of analysis/plotting scripts allowing for a fully reproducible and automated analysis workflow, using \textsc{showyourwork}~\cite{luger_showyourwork_2021}, for this work is available at Ref.~\cite{mccluskey_github_2022} under an MIT license.
The LLZO raw simulation trajectories are available on Zenodo shared under a CC BY-SA 4.0 licence~\cite{coles_llzo_zenodo_2022}.
The method outlined in this work is implemented in the open-source Python package \textsc{kinisi}~\cite{mccluskey_kinisi_2022}, which is available under an MIT license.
